---
title: "strings_factors"
author: "Hana Akbarnejad"
date: "10/15/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(rvest)
library(p8105.datasets)

```

## string manipulation

```{r}

string_vec = c("my", "name", "is", "jeff")
string_vec

str_detect(string_vec, "jeff")
str_detect(string_vec, "m")

str_replace(string_vec, "jeff","Jeff")
str_replace(string_vec, "m", "M")
```

using str, there are a lot of str stuff in R. everything in stringr package start with str_. 
str_detect(): tell me whether of not this string exists in this vetor. gives T or F.
str_replace(): 

```{r}

string_vec = c(
  "i think we all rule for participating",
  "i think i have been caught",
  "i think this will be quite fun actually",
  "it will be fun, i think"
  )

str_detect(string_vec, "i think$")
str_detect(string_vec, "^i think")
```

^ starts with and ends with $

```{r}

string_vec = c(
  "Y'all remember Pres. HW Bush?",
  "I saw a green bush",
  "BBQ and Bushwalking at Molonglo Gorge",
  "BUSH -- LIVE IN CONCERT!!"
  )

str_detect(string_vec, "[Bb]ush")
```

 strings which contain this or this [Bb]
 
```{r}

string_vec = c(
  '7th inning stretch',
  '1st half soon to begin. Texas won the toss.',
  'she is 5 feet 4 inches tall',
  '3AM - cant sleep :('
  )

str_detect(string_vec, "[0-9][a-zA-Z]")
```
 
## patterns

```{r}
string_vec = c(
  'Its 7:11 in the evening',
  'want to go to 7-11?',
  'my flight is AA711',
  'NetBios: scanning ip 203.167.114.66'
  )

str_detect(string_vec, "7.11")
```

dot(.) matches characters, placeholders and stuff

```{r}
string_vec = c(
  'The CI is [2, 5]',
  ':-]',
  ':-[',
  'I found the answer on pages [6-7]'
  )

str_detect(string_vec, "\\[")
```


```{r}

pulse_data = 
  haven::read_sas("./data/public_pulse_data.sas7bdat") %>%
  janitor::clean_names() %>% 
  pivot_longer(
    bdi_score_bl:bdi_score_12m,
    names_to = "visit", 
    names_prefix = "bdi_score_",
    values_to = "bdi") %>%
  select(id, visit, everything()) %>% 
   mutate(
    visit = str_replace(visit, "bl", "00m"),
    visit = fct_relevel(visit, str_c(c("00", "01", "06", "12"), "m"))) %>%
  arrange(id, visit)
```

str_replace() we can find patterns or "exav=ct match" with only "exact match".

```{r}

nsduh_url = "http://samhda.s3-us-gov-west-1.amazonaws.com/s3fs-public/field-uploads/2k15StateFiles/NSDUHsaeShortTermCHG2015.htm"

table_marj = 
  read_html(nsduh_url) %>% 
  html_nodes(css = "table") %>% 
  .[[1]] %>%
  html_table() %>%
  slice(-1) %>%
  as_tibble()
```

```{r}

data_marj = table_marj %>% 
  select(-contains("P Value")) %>% 
pivot_longer(
  -State,
  names_to = "age_year",
  values_to = "percent"
) %>% 
  separate(age_year, into = c("age", "year"), sep = "\\(") %>% 
  mutate(
    year = str_replace(year, "\\)", ""),
    percent = str_replace(percent, "[a-c]$", ""),
    percent = as.numeric(percent)
  )
```

```{r}

data_marj %>%
  filter(age == "12-17") %>% 
  mutate(State = fct_reorder(State, percent)) %>% 
  ggplot(aes(x = State, y = percent, color = year)) + 
    geom_point() + 
    theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

you should also pay attention to the levels of factor variable and what is going on underneath variables!
releveling takes the factor variable and changes it fundamentally!

## revisiting previous examples
```{r}

weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728", "USC00519397", "USS0023B17S"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY", 
                      USC00519397 = "Waikiki_HA",
                      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())

weather_df %>%
  mutate(name = fct_relevel(name, c("Waikiki_HA", "CentralPark_NY", "Waterhole_WA"))) %>% 
  ggplot(aes(x = name, y = tmax)) + 
  geom_violin(aes(fill = name), color = "blue", alpha = .5) + 
  theme(legend.position = "bottom")
```

instead of relevel, we use reorder with respect to somethhing! here: median tmax

```{r}
weather_df %>%
  mutate(name = fct_reorder(name, tmax)) %>% 
  ggplot(aes(x = name, y = tmax)) + 
  geom_violin(aes(fill = name), color = "blue", alpha = .5) + 
  theme(legend.position = "bottom")
```

airbnb dataset
factor_reorder() is great for making plots readable

```{r}

data("nyc_airbnb")

nyc_airbnb %>%
  filter(neighbourhood_group == "Manhattan") %>% 
  mutate(
    neighbourhood = fct_reorder(neighbourhood, price, na.rm = TRUE)) %>% 
  ggplot(aes(x = neighbourhood, y = price)) +
  geom_boxplot() +
  coord_flip() + 
  ylim(0, 1000)
```

coord_flip(): changing stuff on x and y axis (just representations)

```{r}
data("rest_inspec")

rest_inspec %>% 
  group_by(boro, grade) %>% 
  summarize(n = n()) %>% 
  spread(key = grade, value = n)

rest_inspec =
  rest_inspec %>%
  filter(grade %in% c("A", "B", "C"), boro != "Missing") %>% 
  mutate(boro = str_to_title(boro))

rest_inspec %>% 
  filter(str_detect(dba, "Pizza")) %>% 
  group_by(boro, grade) %>% 
  summarize(n = n()) %>% 
  spread(key = grade, value = n)

rest_inspec %>% 
  filter(str_detect(dba, "[Pp][Ii][Zz][Zz][Aa]")) %>% 
  group_by(boro, grade) %>% 
  summarize(n = n()) %>% 
  spread(key = grade, value = n)

rest_inspec %>% 
  filter(str_detect(dba, "[Pp][Ii][Zz][Zz][Aa]")) %>%
  ggplot(aes(x = boro, fill = grade)) + 
  geom_bar()

rest_inspec %>% 
  filter(str_detect(dba, "[Pp][Ii][Zz][Zz][Aa]")) %>%
  mutate(boro = fct_infreq(boro)) %>%
  ggplot(aes(x = boro, fill = grade)) + 
  geom_bar() 

rest_inspec %>% 
  filter(str_detect(dba, "[Pp][Ii][Zz][Zz][Aa]")) %>%
  mutate(
    boro = fct_infreq(boro),
    boro = replace(boro, which(boro == "Brooklyn"), "HipsterHQ")) %>%
  ggplot(aes(x = boro, fill = grade)) + 
  geom_bar() + 
  scale_fill_viridis(discrete = TRUE)

rest_inspec %>% 
  filter(str_detect(dba, "[Pp][Ii][Zz][Zz][Aa]")) %>%
  mutate(
    boro = fct_infreq(boro),
    boro = str_replace(boro, "Brooklyn", "HipsterHQ")) %>%
  ggplot(aes(x = boro, fill = grade)) + 
  geom_bar()

rest_inspec %>% 
  filter(str_detect(dba, regex("pizza", ignore_case = TRUE))) %>%
  mutate(
    boro = fct_infreq(boro),
    boro = fct_recode(boro, "HipsterHQ" = "Brooklyn")) %>%
  ggplot(aes(x = boro, fill = grade)) + 
  geom_bar() + 
  scale_fill_viridis(discrete = TRUE)


```

